---
title: "ASSIGNMENT 1"
author: "AJITH RAJ PERIYASAMY"
date: "2024-02-04"
output:
  html_document: default
  pdf_document: default
---
# SUMMARY

## NETFLIX USERBASE

### The dataset for the assignment has been sourced from: 
#### https://www.kaggle.com/datasets/arnavsmayan/netflix-userbase-dataset



#### The below report depicts the netflix userbase taking into consideration various factors such as subscription type, monthly revenue generated per subscribers, age, gender, etc.


##  IMPORTING DATASET
#### The first code illustrates how to import a dataset from web to into 'R' studio.
```{r}
Netflix_Userbase<-read.csv("./Netflix Userbase.csv")
head(Netflix_Userbase)
```

## DESCRIPTIVE STATISTICS FOR QUANTITATIVE AND CATEGORICAL VARIABLES

### QUANTITATIVE VARIABLE
#### In this case it can be inferred that the average monthly revenue generated by netflix through subscription is 12.51.

### CATEGORICAL VARIABLE
#### Here I have assumed age to be the categorical variable and it is evident that the average age of netflix subscribers is 38.8.
```{r}
summary(Netflix_Userbase[,3])#QUANTITATIVE VARIABLE

summary(Netflix_Userbase[,7])#CATEGORICAL VARIABLE
```
##  TRANSFORMING VARIABLE
#### Below I have made an attempt to convert integer datatype elements to character datatype.
```{r}
as.character(Netflix_Userbase[,3])
```

##  PLOTTING QUANTITATIVE VARIABLE & SCATTERPLOT

#### The first case shows scatterplots which compares two values i.e, age and monthly revenue.
#### In the second scatterplot I have just taken first ten items and plotted a graph to make it clear and presentable, which clearly establishes the relationship between subscribers age and the revenue generated.
```{r}
plot(Netflix_Userbase$Age,Netflix_Userbase$Monthly.Revenue)

plot(Netflix_Userbase$Age[1:10],Netflix_Userbase$Monthly.Revenue[1:10])

# 

```
##  BAR GRAPH

#### The bar graph below has the type of plans in the X-Axis i.e, Basic, Premium, and Standard and the number of subscibers (Both male and female) in the Y-Axis. It can be inferred that many people have subscribed for the Basic plan than the Premium and the Standard plans.

```{r}
library(ggplot2)

ggplot(Netflix_Userbase) +
 aes(x = Subscription.Type, fill = Gender) +
 geom_bar() +
 scale_fill_hue(direction = 1) +
 labs(y = "No.of Male and Female") +
 theme_minimal()
```
