---
title: "UNIVERSAL BANK ASSIGNMENT"
output:
  html_document: default
  pdf_document: default
---
## PROBLEM STATEMENT:

Universal Bank wants to determine if a new customer will accept a loan offer or not.


***
```{r}
# SETTING UP WORKING DIRECTORY:
setwd("/Users/ajithrajperiyasamy/Desktop/FILES/ASSIGNMENTS/FML/FEB 25TH DEADLINE")
```

```{r}
# CALLING THE REQUIRED LIBRARIES:
library(class)
library(caret)
library(e1071)
library(caTools)
library(dplyr)
```


```{r }
# LOADING UNIVERSAL BANK DATASET
universal_bank <- read.csv("/Users/ajithrajperiyasamy/Desktop/FILES/ASSIGNMENTS/FML/FEB 25TH DEADLINE/UniversalBank.csv")
```
```{r}
# TRANSPOSING THE DATASET:
t(t(names(universal_bank))) 
```
#### The dataset has been transposed for convenient working.

```{r}
# REMOVING ID & ZIP code
universal_bank <- universal_bank[,-c(1,5)]
head(universal_bank)
```
#### Removed ID & ZIP Code columns from the dataset.
```{r}
# CONVERTING EDUCATION INTO FACTOR:
universal_bank$Education <- as.factor(universal_bank$Education)
levels(universal_bank$Education)
```

```{r}
# CONVERTING EDUCATION INTO DUMMY VARIABLE:
dummy_groups <- dummyVars(~., data = universal_bank)
universalbank_new<- as.data.frame(predict(dummy_groups,universal_bank))
head(universalbank_new)
```
#### Transformed categorical predictors with more than two categories(Education) into dummy variables first. 

```{r}
# PREPARING SAMPLES:
set.seed(1)  
train_data <-sample(row.names(universalbank_new),0.6*dim(universalbank_new)[1])
validation_data <- setdiff(row.names(universalbank_new), train_data)  
train.df <- universalbank_new[train_data,]
validation.df <- universalbank_new[validation_data,]
t(t(names(train.df)))
```
```{r}
# PARTITIONING THE DATA INTO TRAINING & VALIDATION SETS:
set.seed(1)
split_data <- sample.split(universalbank_new, SplitRatio = 0.6)
training_dataset <- subset(universalbank_new, split_data == TRUE)
validation_dataset <- subset(universalbank_new, split_data == FALSE)

# PRINTING TRAINING & VALIDATION SETS PARTITION:
print(paste("TRAINING SET:", nrow(training_dataset)))
print(paste("VALIDATION SET:", nrow(validation_dataset)))
```
#### Partitioning the data into Training & Validation sets.

```{r}
# NORMALIZING THE DATA:
train.normalized.df <- train.df[,-10]
valid.normalized.df <- validation.df[,-10]
normalized.values <- preProcess(train.df[, -10], method=c("center", "scale"))
train.normalized.df <- predict(normalized.values, train.df[, -10])
validation.normalized.df <- predict(normalized.values, validation.df[, -10])
```
#### Normalized the values.

### Questions

1. Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code using k = 1. Remember to transform categorical predictors with more than two categories into dummy variables first. Specify the success class as 1 (loan acceptance), and use the default cutoff value of 0.5. How would this customer be classified?

```{r}
# CREATING A NEW CUSTOMER SAMPLE
new_customer <- data.frame(Age = 40, 
                           Experience = 10,
                           Income = 84,
                           Family = 2,
                           CCAvg = 2,
                           Education.1 = 0,
                           Education.2 = 1,
                           Education.3 = 0,
                           Mortgage = 0,
                           Securities.Account = 0,
                           CD.Account = 0,
                           Online = 1,
                           CreditCard = 1)

# NORMALIZING THE NEW CUSTOMER:
new_customer.normalized<- new_customer
new_customer.normalized <- predict(normalized.values, new_customer.normalized)
```

```{r}
# RUNNING PREDICTIONS USING kNN
knn.prediction1 <- class::knn(train = train.normalized.df, 
                       test = new_customer.normalized, 
                       cl = train.df$Personal.Loan, k = 1)
knn.prediction1
```

***

2. What is a choice of k that balances between overfitting and ignoring the predictor information?

```{r}
# CALCULATING ACCURACY FOR EACH VALUE OF k
accuracy.df <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))
for(i in 1:15) {
  knn.prediction1 <- class::knn(train = train.normalized.df, 
                         test = validation.normalized.df, 
                         cl = train.df$Personal.Loan, k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.prediction1, as.factor(validation.df$Personal.Loan),positive = "1")$overall[1]
}

which(accuracy.df[,2] == max(accuracy.df[,2])) 

plot(accuracy.df$k,accuracy.df$overallaccuracy)

```
3. CONFUSION MATRIX FOR VALIDATION DATA USING BEST k
```{r}
# CONFUSION MATRIX FOR VALIDATION DATA USING BEST k.

knn.prediction2 <- class::knn(train = train.normalized.df, 
                        test = validation.normalized.df, 
                        cl = train.df$Personal.Loan, k = 3)

confusionMatrix(knn.prediction2,as.factor(validation.df$Personal.Loan))

```
4. Consider the following customer: Age = 40, Experience = 10, Income = 84,
Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0,
Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit
Card = 1. Classify the customer using the best k.

```{r}
# SEGREGATING CUSTOMER USING BEST K.
new_customer2 = data.frame(Age = 40, 
                           Experience = 10, 
                           Income = 84, 
                           Family = 2,
                           CCAvg = 2, 
                           Education.1 = 0, 
                           Education.2 = 1, 
                           Education.3 = 0, 
                           Mortgage = 0, 
                           Securities.Account = 0, 
                           CD.Account = 0, 
                           Online = 1, 
                           CreditCard = 1)

knn.prediction3 <- class::knn(train = train.normalized.df, 
                         test = new_customer2, 
                         cl = train.df$Personal.Loan, k = 3)
knn.prediction3

# It can be inferred that the customer has been approved for personal loan

```
5a.Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply the k-NN method with the k chosen above. Compare the confusion matrix of the test set with that of the training and validation sets. Comment on the differences and their reason.

```{r}
set.seed(2)

# 50% OF THE DATA INTO TRAINING DATA: 
train.df2 = sample(row.names(universalbank_new),0.5*dim(universalbank_new)[1])

# 30% OF THE DATA INTO VALIDATION DATA:  
validation.df2 = sample(setdiff(row.names(universalbank_new), train.df2), 0.3*dim(universalbank_new)[1])

# 20% OF THE DATA INTO TEST DATA:
test.df2 = setdiff(row.names(universalbank_new),union(train.df2,validation.df2))

train.normalized.df2 = universalbank_new[train.df2,]
validation.normalized.df2 = universalbank_new[validation.df2,]
test.normalized.df2 = universalbank_new[test.df2,]

# TRANSPOSING THE DATA:
t(t(names(train.normalized.df2)))


# USING k-NN WITH THE CHOSEN K:

train_knn2 = knn(train = train.normalized.df2[,-8], test = train.normalized.df2[,-8], cl = train.normalized.df2[,8], k=3)

validation_knn2 = knn(train = train.normalized.df2[,-8], test = validation.normalized.df2[,-8], cl = train.normalized.df2[,8], k=3)

test_knn2 = knn(train = train.normalized.df2[,-8], test = test.normalized.df2[,-8], cl = train.normalized.df2[,8], k=3)
```
# COMPARING CONFUSION MATRIX OF TRAINING, VALIDATION & TESTING SETS:
```{r}

confusionmatrixtrain_knn2 = confusionMatrix(train_knn2, as.factor(train.normalized.df2$Personal.Loan),positive = "1")

confusionmatrixtrain_knn2


confusionmatrixvalidation_knn2 = confusionMatrix(validation_knn2, as.factor(validation.normalized.df2$Personal.Loan),positive = "1")

confusionmatrixvalidation_knn2


confusionmatrixtest_knn2 = confusionMatrix(test_knn2, as.factor(test.normalized.df2$Personal.Loan),positive = "1")

confusionmatrixtrain_knn2
```
#### 5b. Compare the confusion matrix of the test set with that of the training and validation sets. Comment on the differences and their reason.
####  The variations in the confusion matrices of the test set compared to the training and validation sets happens due to multiple reasons like Overfitting, Data Variability, Sample Size, Randomness.